Proof nets for first-order additive linear logic

### 2019-03-29 5 - FSCD 2019 - Submitted/Reviews/rebuttal v1.txt

(C) Is the equivalence generated by unification a form of observational equivalence?

We believe, yes, but none of us is an expert.

(D) Both proofs and witness nets are not considered up to alpha-equivalence.

In fact, we may alpha-convert at the level of proofs and proof nets; but for the sanity of the translations between them, not at the level of individual formulas in a sequent proof.

Normally in proof nets, indeed, every universal instantiation should have a unique eigenvariable. But here we only need this property for each slice individually.

@@

Reviews 2 and 3:

(E) Can this be extended to MALL1?

The situation is mixed. There are two ways forward that are likely to succeed:
 - combine witness nets and MALL slice nets [Hughes & Van
   Glabbeek 2005], using a slicing criterion;
 - combine unification nets and MALL conflict nets [Hughes &
   Heijltjes 2016], using a coalescence criterion.
For both, there are still significant combinatorial challenges (e.g. because coalescence for MALL is not a straightforward extension of that for ALL). Other combinations are unlikely to work, or the groundwork is still incomplete: slice nets do not support coalescence (this looks fundamentally problematic), and conflict nets do not yet have a slicing criterion (work in progress).

Review 3:

(F) Unification seems to break the variable naming convention.

Indeed - a naive unification algorithm would. We should stipulate explicitly that an "initial witness map" (the MGU for a link) uses fresh variables in its range, distinct from any existentially quantified ones (which occur in its domain); then all is well.

### 2019-04-16 7 - FSCD 2019 - Final/ALL1.tex
reviews, rebuttal, reorganise
prep for final submission

@@

%\theoremstyle{plain}
%\newtheorem{proposition}[theorem]{Proposition}

@@

@@

First-order terms and the formulas of first-order $\all$ are generated by the following grammars.
%
\setMidspace{5pt}
\[
\begin{array}{@{}l@{}l}
	t &\Coloneqq x \Mid f(t_1,\dots,t_n)
\\[10pt]
	a &\Coloneqq P(t_1,\dots,t_n) \Mid \dual P(t_1,\dots,t_n)
\\[10pt]
	A &\Coloneqq a \Mid A\+A \Mid A\*A \Mid \exists x.A \Mid \forall x.A
\end{array}
\]
%
Negation $(\dual{\,\cdot\,})$ is applied to predicate symbols, $\dual P$ as a matter of convenience. The \defn{dual} $\dual A$ of an arbitrary formula $A$ is given by DeMorgan. We use the following notational conventions.
%
\[
\begin{tabular}{@{}lll@{}} 
	$x,y,z$ & $\in \var$		& first-order variables\\
	$f,g,h$ & $\in \Sigma_f$	& $n$-ary $(n\geq 0)$ function symbols from a fixed alphabet $\Sigma_f$\\
	$P,Q,R$ & $\in \Sigma_p$	& $n$-ary $(n\geq 0)$ predicate symbols from a fixed alphabet $\Sigma_p$ \\
	$s,t,u$ & $\in \terms$ 		& first-order terms over $\var$ and $\Sigma_f$ \\
	$a,b,c$ & $\in \atom$		& atomic propositions \\
	$A,B,C$ & $\in \form$		& \all1 formulas \\
\end{tabular}
\]
%
A \defn{sequent} $\seq AB$ is a pair of formulas $A$ and $B$. A sequent calculus for \all1 is given in Figure~\ref{fig:sequent calculus}, where each rule has a symmetric counterpart for the first formula in the sequent. We write $\prf\pi AB$ for a proof $\pi$ with conclusion sequent $\seq AB$. Two proofs are \defn{equivalent} $\pi\sim\pi'$ if one is obtained from the other by rule permutations (Appendix~\ref{sec:permutations}, Figure~\ref{fig:permutations}).

@@

\begin{definition}
The \defn{de-sequentialization} $[\pi]$ of a sequent proof $\prf\pi AB$ is the witness pre-net $\net{\deseq[\varnothing]\pi AB}AB$ where the function $\deseq-AB$ is defined inductively as follows.
%
\newcommand\XDeseq[4][\sigma]{
 \left[\vcenter{#2}\right]%
 _{\makebox[0pt][l]{$\scriptstyle{#1}$}}%
 ^{\makebox[0pt][l]{$\scriptstyle{#3\,,\,#4}$}}%
 \rule{35pt}{0pt}
}
\begin{align*}
	\XDeseq{\,\infer[\Qrr1]{\seq a{\dual a}}{}}bc &\quad=\quad \{\link bc\}
\\ \\[-5pt]
	\XDeseq{\infer[\Qrr+i]{\seq A {B_1\+B_2}}{\Prf \pi A{B_i}}} {A'} {B_1'\+B'_2} 
	& \quad=\quad 
	\Deseq{\Prf\pi A{B_i}} {A'} {B_i'}
\\ \\[-5pt]
	\XDeseq{\infer[\Qrr*]{\seq A{B\*C}}{
	 \Prf\pi AB && \Prf{\pi'}AC 
	}}{A'}{B'\*C'}
	& \quad=\quad 
	  \Deseq{\Prf\pi AB}{A'}{B'}
	  \cup
	  \Deseq{\Prf{\pi'}AC}{A'}{C'}
\\ \\[-5pt]
	\XDeseq{\infer[\Qrr?t]{\seq A{\exists x.B}}{\Prf\pi A{B[t/x]}}}{A'}{\exists x.B'}
	& \quad=\quad 
	 \Deseq[{\sigma[t/x]}]{\Prf\pi A{B[t/x]}} {A'} {B'}
\\ \\[-5pt]
	\XDeseq{\infer[\Qrr!]
	  {\seq A{\forall x.B}}{\Prf\pi AB}} {A'} {\forall x.B'}
	& \quad=\quad 
	  \Deseq{\Prf\pi AB}{A'}{B'}
\end{align*}
%
%\begin{align*}
%	\XDeseq{\,\infer[\Qrr1]{\seq a{\dual a}}{}}bc &\quad=\quad \{\link bc\}
%\\ \\[-5pt]
%	\XDeseq{\infer[\Qrr+i]{\seq A {B_1\+B_2}}{\prf \pi A{B_i}}} {A'} {B_1'\+B'_2} 
%	& \quad=\quad 
%	\deseq{\prf\pi A{B_i}} {A'} {B_i'}
%\\ \\[-5pt]
%	\XDeseq{\infer[\Qrr*]{\seq A{B\*C}}{
%	 \prf\pi AB && \prf{\pi'}AC 
%	}}{A'}{B'\*C'}
%	& \quad=\quad 
%	  \deseq{\prf\pi AB}{A'}{B'}
%	  \cup
%	  \deseq{\prf{\pi'}AC}{A'}{C'}
%\\ \\[-5pt]
%	\XDeseq{\infer[\Qrr?t]{\seq A{\exists x.B}}{\prf\pi A{B[t/x]}}}{A'}{\exists x.B'}
%	& \quad=\quad 
%	 \deseq[{\sigma[t/x]}]{\prf\pi A{B[t/x]}} {A'} {B'}
%\\ \\[-5pt]
%	\XDeseq{\infer[\Qrr!]
%	  {\seq A{\forall x.B}}{\prf\pi AB}} {A'} {\forall x.B'}
%	& \quad=\quad 
%	  \deseq{\prf\pi AB}{A'}{B'}
%\end{align*}
\end{definition}

@@

For sequentialization, the links in a pre-net will be labelled with a sequent proof. An axiom link will carry an axiom, and each coalescence step introduces one proof rule. Formalizing this, a \defn{proof linking} $\lambda_\Sigma^\Pi$ is a witness linking $\lambda_\Sigma$ with a \defn{proof labelling} $\Pi\colon \lambda\to\proofs$ assigning a sequent proof to each link. We will use and define $\lambda_\Sigma^\Pi$ as a set of \defn{proof links} $\link CD^\pi$, where we require that $\prf\pi{C\sigma}{D\sigma}$, i.e.\ that $\pi$ proves the conclusion $\seq{C\sigma}{D\sigma}$. A \defn{labelled pre-net} $\net{\lambda_\Sigma^\Pi}AB$ is a witness pre-net $\net{\lambda_\Sigma}AB$ with a proof labelling $\Pi$ on $\lambda_\Sigma$. If $\lambda_\Sigma$ is an axiom linking, we assign an \defn{initial proof labelling} $\lambda_\Sigma^\star$ as follows.
\[
	\lambda_\Sigma^\star = \{~\link ab^\pi~\mid~\link ab\in\lambda_\Sigma~,~\pi=~\infer{\seq{a\sigma}{b\sigma}}{}~\}
\]
For correctness we may coalesce a pre-net directly, without constructing a proof. 
%
Then to recap, we have accumulated the following further notational conventions.
%
\[
\begin{tabular}{@{}lcll@{}}
	$\pi,\phi,\psi$		& $\in$ & $\proofs$ & \all1 sequent proofs\\
	$\kappa,\lambda$ 	& $\subset$ & $\form\times\form$ 	& linkings (sets of pairs of formulas)\\
	$\rho,\sigma,\tau$	& $\colon$  & $\var\rightharpoonup\terms$ & witness maps\\
	$\Sigma,\Theta$		& $\colon$  & $\lambda\to\var\rightharpoonup\terms$		& witness labellings on a linking $\lambda$ \\
	$\Pi,\Phi,\Psi$		& $\colon$  & $\lambda\to\proofs$		& proof labellings on a linking $\lambda$
\end{tabular}
\]
\begin{definition}
%
\renewcommand{\+}{{+}}
\renewcommand{\*}{{\times}}
%
\defn{Strict sequentialization} ($\scoal$) is the rewrite relation on labelled pre-nets generated by the following rules, that replace one or two links by another in a pre-net $\net{\lambda_\Sigma^\Pi}AB$ (where $B$ has a subformula $D_1\+D_2$, $D_1\*D_2$, $\exists x.D$, and $\forall x.D$ respectively).
%
\begin{align*}
\tag{$\srr+i$}
	\link C{D_i}^\pi
	& \scoal
	\link C{D_1\+D_2}^\psi
	&& 
 	\psi=~\vc{\infer[\Qrr+i]{\seq{C\sigma}{D_1\sigma\,\+\,D_2\sigma}}  {\Prf \pi {C\sigma} {D_i\sigma}}}
\\ \\[-8pt]
\tag{$\srr*$}
	\left.\begin{array}{@{}l@{\,}}\link C{D_1}^\pi \\ \link C{D_2}^\phi\end{array}\right\}
	& \scoal
	\link C{D_1\*D_2}^\psi
	&&
	\psi=~\vc{\infer[\Qrr*]{\seq{C\sigma}{D_1\sigma\,\*\,D_2\sigma}}  {
      \Prf \pi {C\sigma} {D_1\sigma} & \Prf \phi {C\sigma} {D_2\sigma}    
	}}
\\ \\[-8pt]
\tag{$\srr?$}
	\link CD^\pi
	& \scoal
	\link[\sigma\minus x]C{\exists x.D}^\psi
	\quad{\scriptstyle(x\,\in\,\dom\sigma)}
	&&
	\psi=~\vc{\infer[\Qrr?{\sigma(x)}]{\seq{C(\sigma\minus x)}{\exists x.D(\sigma\minus x)}} {\Prf \pi {C\sigma} {D\sigma} }}	
\\ \\[-8pt]
\tag{$\srr!$}
	\link CD^\pi
	& \scoal
	\link C{\forall x.D}^\psi
	\quad{\scriptstyle\phantom{\minus x}(x\,\notin\,\sigma)}
	&&
	\psi=~\vc{\infer[\Qrr!]{\seq{C\sigma}{\forall x.D\sigma}} {\Prf \pi {C\sigma} {D\sigma}}}
\end{align*}
%
\defn{Strict coalescence} is the same relation on witness pre-nets, ignoring proof labels, illustrated in Figure~\ref{fig:coalescence}. A witness pre-net $\net{\lambda_\Sigma}AB$ \defn{strict-coalesces} if it reduces to $\net{\{\link[\varnothing]AB\}}AB$. It \defn{strongly} strict-coalesces if any coalescence path terminates at $\net{\{\link[\varnothing]AB\}}AB$.
\end{definition}
% ..................................................

@@

% ..................................................
\begin{definition}
An \all1 \defn{witness proof net} or \defn{witness net} is a witness pre-net $\net{\lambda_\Sigma}AB$ with $\lambda_\Sigma$ an axiom linking, that strict-coalesces. It \defn{sequentializes} to a proof $\pi$ if its initial labelling $\net{\lambda_\Sigma^\star}AB$ reduces in $(\scoal)$ to $\net{\{\link[\varnothing]AB^\pi\}}AB$.
\end{definition}
% ..................................................

@@

% ..................................................
\begin{theorem}
\label{thm:proof->net->proof}
For any \all1 proof $\pi$, the witness net $[\pi]$ sequentializes to $\pi$. 
\end{theorem}

%\begin{proof}
%It follows by induction on $\pi$ that if $\lambda_\Sigma=\deseq{\prf\pi AB}{A'}{B'}$ where $A'\sigma=A$ and $B'\sigma=B$, then $\net{\lambda^\star_\Sigma}AB$ reduces in $(\scoal)$ to $\net{\{\link{A'}{B'}^\pi\}}AB$. The statement is the case $\sigma=\varnothing$.
%\end{proof}
% ..................................................

% ..................................................
\begin{theorem}
\label{thm:net->proof->net}
If $\net{\lambda_\Sigma}AB$ sequentializes to $\pi$, then $[\pi]$ is $\net{\lambda_\Sigma}AB$. 
\end{theorem}

%\begin{proof}
%By induction on the sequentialization path 
%$\net{\lambda_\Sigma^\star}AB\scoal^*\net{\{\link[\varnothing]AB^\pi\}}AB$
%it follows that in every pre-net $\net{\kappa_\Theta^\Phi}AB$ on this path, $\lambda_\Sigma$ is equal to the union over the de-sequentialization of all proof labels $\phi$ in $\Phi$:
%\[
%	\lambda_\Sigma=\bigcup~\{\,\deseq\phi CD \mid \link CD^\phi \in \kappa_\Theta^\Phi\,\}~.
%\]
%The statement is then the case $\kappa_\Theta^\Phi=\{\link[\varnothing]AB^\pi\}$.
%\end{proof}
% ..................................................

@@

%\begin{proof}
%From left to right is by inspection of the critical pairs of sequentialization ($\scoal$). From right to left is by inspection of the rule permutations in Figure~\ref{fig:permutations} in Appendix~\ref{sec:permutations}.
%\end{proof}

@@

	\item[Dependency]
We define a \defn{dependency} relation between a universal quantification $\forall x.A$ and an instantiation $B[t/y]$ of $\exists y.B$ where the eigenvariable $x$ occurs free in $t$ (a standard approach to first-order quantification \cite{Miller-1987,Girard-1996,Heijltjes-2010,Mimram-2011}). For each link, we will require this dependency relation to be \emph{acyclic}, which amounts to \emph{slice-wise} first-order correctness. It captures the non-permutability of universal and existential sequent rules due to the \emph{eigenvariable condition} of the former, as below right.
	%\item[]
\[
\vc{
	\infer[\Qrr*]{\Seq{\exists x.A}{B\*C}}{
	 \infer[\Qrr?s]{\Seq{\exists x.A}B}{\Seq{A[s/x]}B}
	 &
	 \infer[\Qrr?t]{\Seq{\exists x.A}C}{\Seq{A[t/x]}C}
   }}
\qquad\qquad
\vc{
	\infer[\Qrr!]{\Seq{\forall x.A}{\exists y.B}}{
	 \infer[\Qrr?t]{\Seq A{\exists y.B}}{\Seq AB[t/y]}
   }}
   \quad\text{where }x\in\fv(t)
\]
\end{description}

%\noindent
%Formalizing this, for a witness linking $\lambda_\Sigma$ and variable $x$, write $\Sigma(x)\subseteq\terms$ for the \defn{witness set} of $x$, which collects the terms assigned to $x$ by the witness maps in $\Sigma$ for every link in $\lambda$:
%\[
%	\Sigma(x) = \{~\sigma(x)~\mid~\link CD\in\lambda_\Sigma~\}~.
%\]

% ..................................................
\begin{definition}[Slice]
A \defn{slice} $S$ of a formula $A$ and a witness map $\sigma$ is a set of pairs $(A',\sigma')$, where $A'\leq A$ and $\sigma'\supseteq\sigma$, given by $S=\{(A,\sigma)\}\cup S'$ where:
%
\begin{itemize}
	\item
\makebox[65pt][l]{If $A = a$}			then $S'=\varnothing$.
	\item
\makebox[65pt][l]{If $A = B\+C$} 		then $S'=S_B\uplus S_C$ with $S_B$ a slice of $B$ and $\sigma$, and $S_C$ one of $C$ and $\sigma$.
	\item
\makebox[65pt][l]{If $A = B\*C$}		then $S'$ is a slice of $B$ and $\sigma$ or a slice of $C$ and $\sigma$.
	\item
\makebox[65pt][l]{If $A = \exists x.B$}	then $S'=\uplus_{t\,\in\,\terms}S_t$ where each $S_t$ is a slice of $B$ and $\sigma[t/x]$.
	\item
\makebox[65pt][l]{If $A = \forall x.B$}	then $S'$ is a slice of $B$ and $\sigma$.
\end{itemize}
%
\noindent
A \defn{slice} of a sequent $\seq AB$ is a set of links
\[
	\{~\link[\sigma\cup\tau]CD~\mid~(C,\sigma)\in S_A~,~(D,\tau)\in S_B~\}
\]
where $S_A$ is a slice of $A$ and $\varnothing$, and $S_B$ a slice of $B$ and $\varnothing$.  A \defn{slice} of a witness pre-net $\net{\lambda_\Sigma} AB$ is the intersection $\lambda_\Sigma\cap S$ of $\lambda_\Sigma$ with a slice $S$ of $\seq AB$.
\end{definition}
% ..................................................

@@

%As in the propositional case, for correctness we will require that each slice is a singleton. In addition, within each slice we need to ensure that the order in which quantifiers are instantiated is sound. That is, we need to have an analogue of the \emph{eigenvariable condition} on the $\qrr!$-rule of sequent calculus, that the quantified variable $x$ of the formula $\forall x.A$ does not occur free in the context. We will define a \defn{dependency} condition to this effect. For simplicity we will define the dependency condition on individual links rather than on slices.

@@

% ..................................................
\begin{definition}
A witness pre-net is \defn{correct} if:
\begin{itemize}
\item it has local eigenvariables and exact coverage,
\item it is \defn{slice-correct}: every slice is a singleton, and
\item it is \defn{dependency-correct}: every column is a partial order (i.e.\ is acyclic/antisymmetric).
\end{itemize}
\end{definition}
% ..................................................

%In the remainder of this section we will establish that the two correctness conditions, by coalescence and by slicing, are equivalent. The geometric condition further gives \emph{strong} coalescence: since it is preserved, no coalescence step yields an incorrect, and thus non-coalescing, pre-net. Finally, canonicity follows by inspecting the critical pairs of sequentialization.

@@

% ..................................................
%\newcounter{lemcorrect}
%\setcounter{lemcorrect}{\thetheorem}
%\begin{lemma}
%\label{lem:correct}
%Strict coalescence preserves and reflects correctness.
%\end{lemma}

%\begin{proof}
%See the appendix.
%\end{proof}

%\newcounter{lemcoalescenceapplies}
%\setcounter{lemcoalescenceapplies}{\thetheorem}
%\begin{lemma}
%\label{lem:coalescence applies}
%To a correct witness pre-net $\net{\lambda_\Sigma}AB$ a coalescence step applies, unless it is fully coalesced already, $\lambda_\Sigma=\{\link[\varnothing]AB\}$.
%\end{lemma}

%\begin{proof}
%See the appendix.
%\end{proof}

@@

@@

%\paragraph*{Acknowledgements}
%
%We would like to thank the anonymous referees for their constructive feedback. Dominic would like to thank his host, Wes Holliday, at the UC Berkeley Logic Group. The first author was supported by EPSRC Project EP/R029121/1.

>>>>

\nolinenumbers

@@

@@

%\relatedversion{http://hal.inria.fr/hal-01867625/}

\EventEditors{Herman Geuvers}
\EventNoEds{1}
\EventLongTitle{4th International Conference on Formal Structures for Computation and Deduction (FSCD 2019)}
\EventShortTitle{FSCD 2019}
\EventAcronym{FSCD}
\EventYear{2019}
\EventDate{June 24--30, 2019}
\EventLocation{Dortmund, Germany}
\EventLogo{}
\SeriesVolume{131}
\ArticleNo{22}

@@

First-order terms and the formulas of first-order $\all$ are generated by the following grammars.
%
\setMidspace{5pt}
\[
\begin{array}{@{}l@{}l}
	t &\Coloneqq x \Mid f(t_1,\dots,t_n)
\\[10pt]
	a &\Coloneqq P(t_1,\dots,t_n) \Mid \dual P(t_1,\dots,t_n)
\\[10pt]
	A &\Coloneqq a \Mid A\+A \Mid A\*A \Mid \exists x.A \Mid \forall x.A
\end{array}
\]
%
Negation $(\dual{\,\cdot\,})$ is applied to predicate symbols, $\dual P$ as a matter of convenience. The \defn{dual} $\dual A$ of an arbitrary formula $A$ is given by DeMorgan. We use the following notational conventions.
%
\[
\begin{tabular}{@{}lll@{}} 
	$x,y,z$ & $\in \var$		& first-order variables\\
	$f,g,h$ & $\in \Sigma_f$	& $n$-ary $(n\geq 0)$ function symbols from a fixed alphabet $\Sigma_f$\\
	$P,Q,R$ & $\in \Sigma_p$	& $n$-ary $(n\geq 0)$ predicate symbols from a fixed alphabet $\Sigma_p$ \\
	$s,t,u$ & $\in \terms$ 		& first-order terms over $\var$ and $\Sigma_f$ \\
	$a,b,c$ & $\in \atom$		& atomic propositions \\
	$A,B,C$ & $\in \form$		& \all1 formulas \\
\end{tabular}
\]
%
A \defn{sequent} $\seq AB$ is a pair of formulas $A$ and $B$. A sequent calculus for \all1 is given in Figure~\ref{fig:sequent calculus}, where each rule has a symmetric counterpart for the first formula in the sequent. We write $\prf\pi AB$ for a proof $\pi$ with conclusion sequent $\seq AB$. Two proofs are \defn{equivalent} $\pi\sim\pi'$ if one is obtained from the other by rule permutations, given in Figure~\ref{fig:permutations}.

@@

\begin{definition}
The \defn{de-sequentialization} $[\pi]$ of a sequent proof $\prf\pi AB$ is the witness pre-net $\net{\deseq[\varnothing]\pi AB}AB$ where the function $\deseq-AB$ is defined inductively as follows (symmetric cases are omitted).
%
\newcommand\XDeseq[4][\sigma]{
 \left[\vcenter{#2}\right]%
 _{\makebox[0pt][l]{$\scriptstyle{#1}$}}%
 ^{\makebox[0pt][l]{$\scriptstyle{#3\,,\,#4}$}}%
 \rule{35pt}{0pt}
}
\begin{align*}
	\XDeseq{\,\infer[\Qrr1]{\seq a{\dual a}}{}}bc &\quad=\quad \{\link bc\}
\\ \\[-5pt]
	\XDeseq{\infer[\Qrr+i]{\seq A {B_1\+B_2}}{\Prf \pi A{B_i}}} {A'} {B_1'\+B'_2} 
	& \quad=\quad 
	\Deseq{\Prf\pi A{B_i}} {A'} {B_i'}
\\ \\[-5pt]
	\XDeseq{\infer[\Qrr*]{\seq A{B\*C}}{
	 \Prf\pi AB && \Prf{\pi'}AC 
	}}{A'}{B'\*C'}
	& \quad=\quad 
	  \Deseq{\Prf\pi AB}{A'}{B'}
	  \cup
	  \Deseq{\Prf{\pi'}AC}{A'}{C'}
\\ \\[-5pt]
	\XDeseq{\infer[\Qrr?t]{\seq A{\exists x.B}}{\Prf\pi A{B[t/x]}}}{A'}{\exists x.B'}
	& \quad=\quad 
	 \Deseq[{\sigma[t/x]}]{\Prf\pi A{B[t/x]}} {A'} {B'}
\\ \\[-5pt]
	\XDeseq{\infer[\Qrr!]
	  {\seq A{\forall x.B}}{\Prf\pi AB}} {A'} {\forall x.B'}
	& \quad=\quad 
	  \Deseq{\Prf\pi AB}{A'}{B'}
\end{align*}
%
%\begin{align*}
%	\XDeseq{\,\infer[\Qrr1]{\seq a{\dual a}}{}}bc &\quad=\quad \{\link bc\}
%\\ \\[-5pt]
%	\XDeseq{\infer[\Qrr+i]{\seq A {B_1\+B_2}}{\prf \pi A{B_i}}} {A'} {B_1'\+B'_2} 
%	& \quad=\quad 
%	\deseq{\prf\pi A{B_i}} {A'} {B_i'}
%\\ \\[-5pt]
%	\XDeseq{\infer[\Qrr*]{\seq A{B\*C}}{
%	 \prf\pi AB && \prf{\pi'}AC 
%	}}{A'}{B'\*C'}
%	& \quad=\quad 
%	  \deseq{\prf\pi AB}{A'}{B'}
%	  \cup
%	  \deseq{\prf{\pi'}AC}{A'}{C'}
%\\ \\[-5pt]
%	\XDeseq{\infer[\Qrr?t]{\seq A{\exists x.B}}{\prf\pi A{B[t/x]}}}{A'}{\exists x.B'}
%	& \quad=\quad 
%	 \deseq[{\sigma[t/x]}]{\prf\pi A{B[t/x]}} {A'} {B'}
%\\ \\[-5pt]
%	\XDeseq{\infer[\Qrr!]
%	  {\seq A{\forall x.B}}{\prf\pi AB}} {A'} {\forall x.B'}
%	& \quad=\quad 
%	  \deseq{\prf\pi AB}{A'}{B'}
%\end{align*}
\end{definition}

@@

For sequentialization, the links in a pre-net will be labelled with a sequent proof. An axiom link will carry an axiom, and each coalescence step introduces one proof rule. Formalizing this, a \defn{proof linking} $\lambda_\Sigma^\Pi$ is a witness linking $\lambda_\Sigma$ with a \defn{proof labelling} $\Pi\colon \lambda\to\proofs$ assigning a sequent proof to each link. We will use and define $\lambda_\Sigma^\Pi$ as a set of \defn{proof links} $\link CD^\pi$, where we require that $\prf\pi{C\sigma}{D\sigma}$, i.e.\ that $\pi$ proves the conclusion $\seq{C\sigma}{D\sigma}$. A \defn{labelled pre-net} $\net{\lambda_\Sigma^\Pi}AB$ is a witness pre-net $\net{\lambda_\Sigma}AB$ with a proof labelling $\Pi$ on $\lambda_\Sigma$. If $\lambda_\Sigma$ is an axiom linking, we assign an \defn{initial proof labelling} $\lambda_\Sigma^\star$ as follows.
\[
	\lambda_\Sigma^\star = \{~\link ab^\pi~\mid~\link ab\in\lambda_\Sigma~,~\pi=~\infer{\seq{a\sigma}{b\sigma}}{}~\}
\]
For correctness we may coalesce a pre-net directly, without constructing a proof. 
%
Then to recap, we have accumulated the following further notational conventions.
%
\[
\begin{tabular}{@{}lcll@{}}
	$\pi,\phi,\psi$		& $\in$ & $\proofs$ & \all1 sequent proofs\\
	$\kappa,\lambda$ 	& $\subset$ & $\form\times\form$ 	& linkings (sets of pairs of formulas)\\
	$\rho,\sigma,\tau$	& $\colon$  & $\var\rightharpoonup\terms$ & witness maps\\
	$\Sigma,\Theta$		& $\colon$  & $\lambda\to\var\rightharpoonup\terms$		& witness labellings on a linking $\lambda$ \\
	$\Pi,\Phi,\Psi$		& $\colon$  & $\lambda\to\proofs$		& proof labellings on a linking $\lambda$
\end{tabular}
\]

\begin{definition}
%
\renewcommand{\+}{{+}}
\renewcommand{\*}{{\times}}
%
\defn{Strict sequentialization} ($\scoal$) is the rewrite relation on labelled pre-nets generated by the following rules, that replace one or two links by another in a pre-net $\net{\lambda_\Sigma^\Pi}AB$ where $B$ has a subformula $D_1\+D_2$, $D_1\*D_2$, $\exists x.D$, and $\forall x.D$ respectively. Symmetric cases are omitted.
%
\begin{align*}
\tag{$\srr+i$}
	\link C{D_i}^\pi
	& \scoal
	\link C{D_1\+D_2}^\psi
	&& 
 	\psi=~\vc{\infer[\Qrr+i]{\seq{C\sigma}{D_1\sigma\,\+\,D_2\sigma}}  {\Prf \pi {C\sigma} {D_i\sigma}}}
\\ \\[-8pt]
\tag{$\srr*$}
	\left.\begin{array}{@{}l@{\,}}\link C{D_1}^\pi \\ \link C{D_2}^\phi\end{array}\right\}
	& \scoal
	\link C{D_1\*D_2}^\psi
	&&
	\psi=~\vc{\infer[\Qrr*]{\seq{C\sigma}{D_1\sigma\,\*\,D_2\sigma}}  {
      \Prf \pi {C\sigma} {D_1\sigma} & \Prf \phi {C\sigma} {D_2\sigma}    
	}}
\\ \\[-8pt]
\tag{$\srr?$}
	\link CD^\pi
	& \scoal
	\link[\sigma\minus x]C{\exists x.D}^\psi
	\quad{\scriptstyle(x\,\in\,\dom\sigma)}
	&&
	\psi=~\vc{\infer[\Qrr?{\sigma(x)}]{\seq{C(\sigma\minus x)}{\exists x.D(\sigma\minus x)}} {\Prf \pi {C\sigma} {D\sigma} }}	
\\ \\[-8pt]
\tag{$\srr!$}
	\link CD^\pi
	& \scoal
	\link C{\forall x.D}^\psi
	\quad{\scriptstyle\phantom{\minus x}(x\,\notin\,\sigma)}
	&&
	\psi=~\vc{\infer[\Qrr!]{\seq{C\sigma}{\forall x.D\sigma}} {\Prf \pi {C\sigma} {D\sigma}}}
\end{align*}
%
\defn{Strict coalescence} is the same relation on witness pre-nets, ignoring proof labels, illustrated in Figure~\ref{fig:coalescence}. A witness pre-net $\net{\lambda_\Sigma}AB$ \defn{strict-coalesces} if it reduces to $\net{\{\link[\varnothing]AB\}}AB$. It \defn{strongly} strict-coalesces if any coalescence path terminates at $\net{\{\link[\varnothing]AB\}}AB$.
\end{definition}

@@

\begin{definition}
An \all1 \defn{witness proof net} or \defn{witness net} is a witness pre-net $\net{\lambda_\Sigma}AB$ with $\lambda_\Sigma$ an axiom linking, that strict-coalesces. It \defn{sequentializes} to a proof $\pi$ if its initial labelling $\net{\lambda_\Sigma^\star}AB$ reduces in $(\scoal)$ to $\net{\{\link[\varnothing]AB^\pi\}}AB$.
\end{definition}

@@

\begin{theorem}
\label{thm:proof->net->proof}
For any \all1 proof $\pi$, the witness net $[\pi]$ sequentializes to $\pi$. 
\end{theorem}

\begin{theorem}
\label{thm:net->proof->net}
If $\net{\lambda_\Sigma}AB$ sequentializes to $\pi$, then $[\pi]$ is $\net{\lambda_\Sigma}AB$. 
\end{theorem}

@@

@@

	\item[Dependency]
We define a \defn{dependency} relation between a universal quantification $\forall x.A$ and an instantiation $B[t/y]$ of $\exists y.B$ where the eigenvariable $x$ occurs free in $t$ (a standard approach to first-order quantification \cite{Miller-1987,Girard-1996,Heijltjes-2010,Mimram-2011}). For each link, we will require this dependency relation to be \emph{acyclic}, which amounts to \emph{slice-wise} first-order correctness. It captures the non-permutability of universal and existential sequent rules due to the \emph{eigenvariable condition} of the former, as below right.
\end{description}

\[
\vc{
	\infer[\Qrr*]{\Seq{\exists x.A}{B\*C}}{
	 \infer[\Qrr?s]{\Seq{\exists x.A}B}{\Seq{A[s/x]}B}
	 &
	 \infer[\Qrr?t]{\Seq{\exists x.A}C}{\Seq{A[t/x]}C}
   }}
\qquad\qquad
\vc{
	\infer[\Qrr!]{\Seq{\forall x.A}{\exists y.B}}{
	 \infer[\Qrr?t]{\Seq A{\exists y.B}}{\Seq AB[t/y]}
   }}
   \quad\text{where }x\in\fv(t)
\]

\noindent
\emph{Expansion} covers the interaction between existential quantifiers and products, and \emph{dependency} that between existential and universal quantifiers. Note that in proof nets that include the multiplicatives (e.g.\ \cite{Girard-1996,Hughes-2018}), instead of a dependency as a \emph{partial order} it is common to consider the underlying graph, created by adding a \emph{jump} or \emph{leap} edge when an existential instantiation $B[t/y]$ of $\exists y.B$ contains the eigenvariable $x$ of $\forall x.A$. These edges participate in the \emph{switching} condition \cite{Danos-Regnier-1989}, to capture the interaction between quantifiers and tensors. Without the tensor, here, the dependency as a partial order suffices.

\begin{definition}[Slice]
A \defn{slice} $S$ of a formula $A$ and a witness map $\sigma$ is a set of pairs $(A',\sigma')$, where $A'\leq A$ and $\sigma'\supseteq\sigma$, given by $S=\{(A,\sigma)\}\cup S'$ where:
%
\begin{itemize}
	\item
\makebox[65pt][l]{If $A = a$}			then $S'=\varnothing$.
	\item
\makebox[65pt][l]{If $A = B\+C$} 		then $S'=S_B\uplus S_C$ with $S_B$ a slice of $B$ and $\sigma$, and $S_C$ one of $C$ and $\sigma$.
	\item
\makebox[65pt][l]{If $A = B\*C$}		then $S'$ is a slice of $B$ and $\sigma$ or a slice of $C$ and $\sigma$.
	\item
\makebox[65pt][l]{If $A = \exists x.B$}	then $S'=\uplus_{t\,\in\,\terms}S_t$ where each $S_t$ is a slice of $B$ and $\sigma[t/x]$.
	\item
\makebox[65pt][l]{If $A = \forall x.B$}	then $S'$ is a slice of $B$ and $\sigma$.
\end{itemize}
%
\noindent
A \defn{slice} of a sequent $\seq AB$ is a set of links
\[
	\{~\link[\sigma\cup\tau]CD~\mid~(C,\sigma)\in S_A~,~(D,\tau)\in S_B~\}
\]
where $S_A$ is a slice of $A$ and $\varnothing$, and $S_B$ a slice of $B$ and $\varnothing$.  A \defn{slice} of a witness pre-net $\net{\lambda_\Sigma} AB$ is the intersection $\lambda_\Sigma\cap S$ of $\lambda_\Sigma$ with a slice $S$ of $\seq AB$.
\end{definition}

@@

@@

\begin{definition}
A witness pre-net is \defn{correct} if:
\begin{itemize}
\item it has local eigenvariables and exact coverage,
\item it is \defn{slice-correct}: every slice is a singleton, and
\item it is \defn{dependency-correct}: every column is a partial order (i.e.\ is acyclic/antisymmetric).
\end{itemize}
\end{definition}

@@

@@

% ..................................................
\begin{figure}[!ht]
\[
\renewcommand\+{{+}}
\renewcommand\*{{\times}}
\scalebox{0.85}{$
\begin{array}{c@{\qquad}c}
	\permAC   & \permEC   
\\	\downperm & \downperm \\
	\permCA   & \permCE
\\ \\[20pt]
	\permPC   & \permXC 
\\	\downperm & \downperm \\
	\permCP   & \permCX
\end{array}
\begin{array}{c}
	\permCD \\ \downperm \\ \permDC
\end{array}
$}
\]
\caption{Cut-permutations}
\label{fig:cut-permutations}
\end{figure}
% ..................................................

@@

\section*{Acknowledgements}
